---
title: "COVID 19 Analysis"
date: ''
output:
  pdf_document: default
  html_document:
    df_print: paged
---

#### Required Packages

```{r setup, include=FALSE}
library(tidyverse)
library(lubridate)
library(usmap)
```

#### Part 1 - Basic Exploration of US Data

The New York Times (the Times) has aggregated reported COVID-19 data from state and local governments and health departments since 2020 and provides public access through a repository on GitHub. One of the data sets provided by the Times is county-level data for cumulative cases and deaths each day. This will be your primary data set for the first two parts of your analysis. 

County-level COVID data from 2020, 2021, and 2022 has been imported below. Each row of data reports the cumulative number of cases and deaths for a specific county each day. A FIPS code, a standard geographic identifier, is also provided which you will use in Part 2 to construct a map visualization at the county level for a state. 

Additionally, county-level population estimates reported by the US Census Bureau has been imported as well. You will use these estimates to caluclate statistics per 100,000 people. 

```{r import-nyt-data}
# Import New York Times COVID-19 data
# Import Population Estimates from US Census Bureau 

us_counties_2020 <- read_csv("https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties-2020.csv")
us_counties_2021 <- read_csv("https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties-2021.csv")
us_counties_2022 <- read_csv("https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties-2022.csv")

us_population_estimates <- read_csv("fips_population_estimates.csv")
```

##### Question 1 

Your first task is to combine and tidy the 2020, 2021, and 2022 COVID data sets and find the total deaths and cases for each day since March 15, 2020 (2020-03-15). The data sets provided from the NY Times also includes statistics from Puerto Rico, a US territory. You may remove these observations from the data as they will not be needed for your analysis. Once you have tidied the data, find the total COVID-19 cases and deaths since March 15, 2020. Write a sentence or two after the code block communicating your results. Use inline code to include the `max_date`, `us_total_cases`, and `us_total_deaths` variables. To write inline code use `r `. 

```{r p1q1-response}

# Combine and tidy the 2020, 2021, and 2022 COVID data sets. 
# Hint: Review the rbind() documentation to combine the three data sets. 
#
## YOUR CODE HERE ##

# combining all 3 datasets to form one df using rbind
combined_df <- rbind(us_counties_2020, us_counties_2021, us_counties_2022)

# performing the required tidying and other data preprocessing of data needed
df1 <- combined_df %>% 
  filter(state != 'Puerto Rico', date >= '2020-03-15') %>% # filtering out the rows with Puerto Rico as state
  arrange(date) %>% # arranging as per date in ascending order
  select(date, cases, deaths) %>% # selecting the only required columns
  group_by(date) %>% # grouping by date to sum up total deaths and cases 
  summarise(total_deaths = sum(deaths), 
            total_cases = sum(cases))


# creating the 3 variables for max date, total deaths and total cases and using the tail function to get the required answer
max_date <- tail(df1, n = 1)$date # replace the quotes with your code to find the most recent date in the data set
us_total_cases <- tail(df1, n = 1)$total_cases
us_total_deaths <- tail(df1, n = 1)$total_deaths
df1
```

```{r p1q1-sample}

# Your output should look similar to the following tibble: 
#
#   A tibble: 657 x 3
#       date          total_deaths   total_cases
#      <date>            <dbl>         <dbl>
#   1 2020-03-15           68          3595
#   2 2020-03-16           91          4502
#   3 2020-03-17          117          5901
#   4 2020-03-18          162          8345
#   5 2020-03-19          212         12387
#   6 2020-03-20          277         17998
#   7 2020-03-21          359         24507
#   8 2020-03-22          457         33050
#   9 2020-03-23          577         43474
#  10 2020-03-24          783         53899
# ... with 647 more rows
#

```

-- Communicate your methodology, results, and interpretation here --

Using the rbind function, it is possible to combine the three datasets as required. The US territory 'Puerto Rico' can be filtered away first and the dataset can be arranged and necessary columns was selected. The dateset was grouped by the date columns since we need the total deaths and cases for every date. And since we already arranged the data in ascending order previous, the last row after grouping ans summing up the deaths and cases columns will give you the required answers for the total deaths and cases as on '2022-12-31'
As of `r format(max_date, "%B %d, %Y")`, the total number of deaths in the US was `r us_total_deaths` arising out of `r us_total_cases` total cases! 
Assuming each person that tested positive for covid, did so only once - we can estimate that the mortality rate was `r (us_total_deaths/us_total_cases)`

##### Question 2 

Create a visualization for the total number of deaths and cases in the US since March 15, 2020. Before you create your visualization, review the types of plots you can create using the ggplot2 library and think about which plots would be effective in communicating your results. After you have created your visualization, write a few sentences describing your visualization. How could the plot be interpreted? Could it be misleading? 

```{r p1q2-response}

# Create a visualization for the total number of US cases and deaths since March 15, 2020. 
#
## YOUR CODE HERE ##

library(scales)

df1$total_deaths <- as.numeric(df1$total_deaths)
df1$total_cases <- as.numeric(df1$total_cases)

coeff <- 90

ggplot(df1, aes(x = date)) +
  geom_bar(aes(y = total_deaths, color = "Cumulative Deaths"), size = 0.1, stat = "identity", alpha = 1) +
  geom_line(aes(y = total_cases/coeff, color = "Cumulative Cases"), size = 1, position = position_nudge(x = -1)) + 
  labs(x = "Date", color = "Metric") +
  scale_color_manual(values = c("Cumulative Deaths" = "red", "Cumulative Cases" = "blue")) +
  theme_bw() +
  scale_y_continuous(
    name = "Cumulative Cases",
    labels = unit_format(unit = "M", scale = 100e-6), 
    sec.axis = sec_axis(~.*10, name = "Cumulative Deaths", labels = unit_format(unit = "M", scale = 1e-6))  # Adjust the transformation
  ) +
ggtitle("When cases go up, deaths go up too!")
```

-- Communicate your methodology, results, and interpretation here -- 

The total number of US cases and deaths is visualized using line and bar plots. Mainly, making use of a secondary y-axis to have both cases and deaths represented in a single chart. Based on the values of total deaths and total cases, it is apparent that the cases is around 99 million,  and deaths is just over 1 million. This makes cases a 100 times multiple of deaths approximately. Hence I'm taking a coefficient of 100 to get the scales to be similar to get the visualization better. 
A bar plot is used for deaths which is in red in color and a line plot in blue for total cases. The left y-axis indicates the cumulative cases and the right y-axis which is the secondary axis indicates the cumulative deaths. 
It is apparent that when the total number of cases go up, the number of deaths go up too. 

##### Question 3

While it is important to know the total deaths and cases throughout the COVID-19 pandemic, it is also important for local and state health officials to know the the number of new cases and deaths each day to understand how rapidly the virus is spreading. Using the table you created in Question 1, calculate the number of new deaths and cases each day and a seven-day average of new deaths and cases. Once you have organized your data, find the days that saw the largest number of new cases and deaths. Write a sentence or two after the code block communicating your results.

```{r p1q3-response}

# Create a new table, based on the table from Question 1, and calculate the number of new deaths and cases each day and a seven day average of new deaths and cases. 
#
# Hint: Look at the documentation for lag() when computing the number of new deaths and cases and the seven-day averages. 
#
#
## YOUR CODE HERE ##

# loading the zoo library needed to calculate the rolling averages 
library(zoo)

# calculating the daily changes in cases and deaths and the 7 day rolling averages for both 
df2 <- df1 %>% 
  mutate(delta_deaths_1 = c(NA, diff(total_deaths)), 
         delta_cases_1 = c(NA, diff(total_cases)), 
         delta_deaths_7 = rollapply(delta_deaths_1, width = 7, FUN = mean, align = "right", fill = NA), 
         delta_cases_7 = rollapply(delta_cases_1, width = 7, FUN = mean, align = "right", fill = NA))

# viewing the new df
df2

# getting the dates which had the maximum occurrences of cases and deaths 
max_new_cases_date <- df2[which.max(df2$delta_cases_1), ]$date
max_new_deaths_date <- df2[which.max(df2$delta_deaths_1), ]$date
min_new_cases_date <- df2[which.min(df2$delta_cases_1), ]$date
min_new_deaths_date <- df2[which.min(df2$delta_deaths_1), ]$date

```

```{r p1q3-sample}

#  Your output should look similar to the following tibble: 
#  
#  date
#  total_deaths     >  the cumulative number of deaths up to and including the associated date
#  total_cases      >  the cumulative number of cases up to and including the associated date
#  delta_deaths_1   >  the number of new deaths since the previous day
#  delta_cases_1    >  the number of new cases since the previous day
#  delta_deaths_7   >  the average number of deaths in a seven-day period
#  delta_cases_7    >  the average number of cases in a seven-day period
#==
#  A tibble: 813 x 7
#    date           total_deaths   total_cases   delta_deaths_1    delta_cases_1  delta_deaths_7  delta_cases_7
#    <date>            <dbl>          <dbl>         <dbl>               <dbl>        <dbl>            <dbl>
#  1 2020-03-15           68          3600            0                     0         NA                NA 
#  2 2020-03-16           91          4507           23                   907         NA                NA 
#  3 2020-03-17          117          5906           26                  1399         NA                NA 
#  4 2020-03-18          162          8350           45                  2444         NA                NA 
#  5 2020-03-19          212         12393           50                  4043         NA                NA 
#  6 2020-03-20          277         18012           65                  5619         NA                NA 
#  7 2020-03-21          360         24528           83                  6516         NA                NA 
#  8 2020-03-22          458         33073           98                  8545       55.7              4210.
#  9 2020-03-23          579         43505          121                 10432       69.7              5571.
# 10 2020-03-24          785         53938          206                 10433       95.4              6862.
# ... with 803 more rows



```

-- Communicate your methodology, results, and interpretation here -- 

Used the function rollapply to get the 7 day rolling averages and diff function to the get the daily changes in cases and deaths. 
The date that had the maximum no of cases is - `r df2[which.max(df2$delta_cases_1), ]$date`
The date that had the maximum no of deaths is - `r df2[which.max(df2$delta_deaths_1), ]$date`
The date that had the minimum no of cases is - `r df2[which.min(df2$delta_cases_1), ]$date`
The date that had the minimum no of deaths is - `r df2[which.min(df2$delta_deaths_1), ]$date`

##### Question 4


```{r p1q4-response}

# Create a new table, based on the table from Question 3, and calculate the number of new deaths and cases per 100,000 people each day and a seven day average of new deaths and cases per 100,000 people. 

# Hint: To calculate per 100,000 people, first tidy the population estimates data and calculate the US population in 2020 and 2021. Then, you will need to divide  each statistic by the estimated population and then multiply by 100,000.
#
# Hint: look at the help documentation for grepl() and case_when() to divide the averages by the US population for each year. 
# For example, take the simple tibble, t_new: 
#
#     x     y   
#   <int> <chr>  
#     1     a   
#     2     b   
#     3     a    
#     4     b  
#     5     a  
#     6     b
#
#
# To add a column, z, that is dependent on the value in y, you could: 
#
# t_new %>%  
#   mutate(z = case_when(grepl("a", y) ~ "not b", 
#                        grepl("b", y) ~ "not a"))
#

## YOUR CODE HERE ##

# calculating the total population of the US
pop_df <- us_population_estimates %>% 
  group_by(Year) %>% 
  summarise(total_pop = sum(Estimate))

# creating separate variables to hold these population numbers 
pop_2020 <- pop_df$total_pop[1]
pop_2021 <- pop_df$total_pop[2]

# using the grepl and case_when functions to get the number of cases and deaths per 100,000 people
# since the us_population_estimates did not have population data for the year 2022, any dates in the year 2022 have been filtered 
df3 <- df2 %>% 
  filter(date < "2022-01-01") %>% 
  mutate(total_deaths = case_when(grepl("2020", date) ~ (total_deaths/pop_2020)*100000, 
                                  grepl("2021", date) ~ (total_deaths/pop_2021)*100000), 
         total_cases = case_when(grepl("2020", date) ~ (total_cases/pop_2020)*100000, 
                                 grepl("2021", date) ~ (total_cases/pop_2021)*100000),
         delta_deaths_1 = case_when(grepl("2020", date) ~ (delta_deaths_1/pop_2020)*100000, 
                                    grepl("2021", date) ~ (delta_deaths_1/pop_2021)*100000),
         delta_cases_1 = case_when(grepl("2020", date) ~ (delta_cases_1/pop_2020)*100000, 
                                   grepl("2021", date) ~ (delta_cases_1/pop_2021)*100000), 
         delta_deaths_7 = case_when(grepl("2020", date) ~ (delta_deaths_7/pop_2020)*100000, 
                                    grepl("2021", date) ~ (delta_deaths_7/pop_2021)*100000),
         delta_cases_7 = case_when(grepl("2020", date) ~ (delta_cases_7/pop_2020)*100000, 
                                   grepl("2021", date) ~ (delta_cases_7/pop_2021)*100000))

# viewing the end data
df3

tail(df3)

summary(df3)
```

```{r p1q4-sample}

#  Your output should look similar to the following tibble: 
#  
#  date
#  total_deaths     >  the cumulative number of deaths up to and including the associated date
#  total_cases      >  the cumulative number of cases up to and including the associated date
#  delta_deaths_1   >  the number of new deaths since the previous day
#  delta_cases_1    >  the number of new cases since the previous day
#  delta_deaths_7   >  the average number of deaths in a seven-day period
#  delta_cases_7    >  the average number of cases in a seven-day period
#==
#  A tibble: 657 x 7
#       date         total_deaths   total_cases   delta_deaths_1   delta_cases_1 delta_deaths_7 delta_cases_7
#      <date>            <dbl>         <dbl>          <dbl>           <dbl>           <dbl>         <dbl>
#   1 2020-03-15         0.0205        1.08                0              0              NA           NA   
#   2 2020-03-16         0.0275        1.36          0.00694          0.274              NA           NA   
#   3 2020-03-17         0.0353        1.78          0.00784          0.422              NA           NA   
#   4 2020-03-18         0.0489        2.52           0.0136          0.737              NA           NA   
#   5 2020-03-19         0.0640        3.74           0.0151           1.22              NA           NA   
#   6 2020-03-20         0.0836        5.43           0.0196           1.69              NA           NA   
#   7 2020-03-21         0.108         7.39           0.0247           1.96              NA           NA   
#   8 2020-03-22         0.138         9.97           0.0296           2.58          0.0168         1.27
#   9 2020-03-23         0.174         13.1           0.0362           3.14          0.0209         1.68
#  10 2020-03-24         0.236         16.3           0.0621           3.14          0.0287         2.07
```

-- Communicate your methodology, results, and interpretation here -- 

When comparing the beginning of 2020 with end of 2021, the number of cases increased from 1 to 16000 roughly per 100,000 people. And the number of deaths also increased from 0.02 to 247 per 100,000 people during the same time period. 

#### Question 5

```{r p1q5-response}

# Create a visualization to compare the seven-day average cases and deaths per 100,000 people. 

# visualization for deaths 
ggplot(df3, aes(x = date)) +
  geom_col(aes(y = delta_deaths_1), fill = "gray80", width = 2, stat = "identity", alpha = 1) +  # Raw scores as bar chart
  geom_line(aes(y = delta_deaths_7), color = "blue", size = 0.8) +  # Rolling average as line
  labs(x = "Date", y = "Daily Deaths", title = "7-day Rolling Average vs Daily Deaths") +
  theme_bw()

# visualization for cases
ggplot(df3, aes(x = date)) +
  geom_col(aes(y = delta_cases_1), fill = "gray80", width = 2, stat = "identity", alpha = 1) +  # Raw scores as bar chart
  geom_line(aes(y = delta_cases_7), color = "blue", size = 0.8) +  # Rolling average as line
  labs(x = "Date", y = "Daily Cases", title = "7-day Rolling Average vs Daily Cases") +
  theme_bw()
```

-- Communicate your methodology, results, and interpretation here -- 

To get a better view of the 7 day rolling averages that is indicated by the blue line, it is compared with the actual raw deaths and cases that occurred on a per day basis that is shown by the gray column bars. There are two visualizations for cases and deaths here. 
It looks like the maximum daily deaths occurred towards the end of 2020 and the beginning of 2021. 
And wrt cases, daily occurrences peaked towards the end of 2020 and a similar trend occurred in the month of Aug-sept and had maximum cases towards the beginning of 2021. 

##### Part 2 - US State Comparison

While understanding the trends on a national level can be helpful in understanding how COVID-19 impacted the United States, it is important to remember that the virus arrived in the United States at different times. For the next part of your analysis, you will begin to look at COVID related deaths and cases at the state and county-levels.

##### Question 1

Your first task in Part 2 is to determine the top 10 states in terms of total deaths and cases between March 15, 2020, and December 31, 2021.

Once you have both lists, briefly describe your methodology and your results.

```{r p2q1-response}

# Determine the top 10 states in terms of total deaths and cases between March 15, 2020, and December 31, 2021. To do this, transform your combined COVID-19 data to summarize total deaths and cases by state up to December 31, 2021. 

# creating a df grouped by state and date to find the total sum of deaths and cases
# appropriate filters have been added as per the question 
state_df <- combined_df %>% 
  filter(state != 'Puerto Rico', date >= "2020-03-15", date <= "2021-12-31") %>% 
  group_by(state, date) %>% 
  summarise(total_deaths = sum(deaths), 
            total_cases = sum(cases))

# using the above df to sort it based on total cases in descending order and obtaining the top 10 states
state_df <- state_df %>% 
  arrange(desc(total_cases)) %>% 
  distinct(state, .keep_all = TRUE)

state_df
```

```{r p2q1-sample}

# Your transformed data should look similar to the following tibble:  
#
# A tibble: 51 x 4
#     state                  date       total_deaths  total_cases
#     <chr>                 <date>          <dbl>       <dbl>
#  1 California           2021-12-31        76709      5515613
#  2 Texas                2021-12-31        76062      4574881
#  3 Florida              2021-12-31        62504      4166392
#  4 New York             2021-12-31        58993      3473970
#  5 Illinois             2021-12-31        31017      2154058
#  6 Pennsylvania         2021-12-31        36705      2036424
#  7 Ohio                 2021-12-31        29447      2016095
#  8 Georgia              2021-12-31        30283      1798497
#  9 Michigan             2021-12-31        28984      1706355
# 10 North Carolina       2021-12-31        19436      1685504
# ... with 41 more rows
```

-- Communicate your methodology, results, and interpretation here -- 

It looks like California was the state that was affected the most in terms of total cases and deaths during the duration of March 2020 to December 2021. Followed by Texas and Florida taking the 2nd and 3rd position in this ranking. 
It seems a bit expected as well since these states are also very highly populated and the chances of a person testing positive in these states are much higher than the other states. 

##### Question 2

Determine the top 10 states in terms of deaths per 100,000 people and cases per 100,000 people between March 15, 2020, and December 31, 2021.

Once you have both lists, briefly describe your methodology and your results. Do you expect the lists to be different than the one produced in Question 1? Which method, total or per 100,000 people, is a better method for reporting the statistics? 

```{r p2q2-response}

# Determine the top 10 states in terms of deaths and cases per 100,000 people between March 15, 2020, and December 31, 2021. You should first tidy and transform the population estimates to include population totals by state. Use your relational data verbs (e.g. full_join()) to join the population estimates with the cases and death statistics using the state name as a key. Then, use case_when() and grepl() to add a population column to your table that only includes the estimated population for the associated year. Finally, mutate your table to calculate deaths and cases per 100,000 people and summarize by state. 


# obtaining the year from the overall date 
state_df <- state_df %>% 
    mutate(year = year(date))

# grouping the population data by state and year to find the sum of total population for each state and year
state_pop_df <- us_population_estimates %>% 
    group_by(STNAME, Year) %>% 
    summarise(total_pop = sum(Estimate)) %>% 
    arrange(STNAME, Year)

# finding the deaths and cases per 100,000 people for each state using its own respective population
state_per_100k <- state_df %>% 
  full_join(state_pop_df, by = c("state" = "STNAME", "year" = "Year")) %>% 
  mutate(deaths_per_100k = (total_deaths/total_pop)*100000, 
         cases_per_100k = (total_cases/total_pop)*100000) %>% 
  select(state, date, deaths_per_100k, cases_per_100k) %>% 
  arrange(desc(cases_per_100k)) %>% 
  head(n = 10)

# viewing the data 
state_per_100k
```

```{r p2q2-sample}

# Your transformed data should look similar to the following tibble:
#
# A tibble: 51 x 4
#     state                   date           deaths_per_100k  cases_per_100k
#     <chr>                  <date>               <dbl>          <dbl>
#  1 North Dakota          2021-12-31             265.           22482.
#  2 Alaska                2021-12-31             130.           21310.
#  3 Rhode Island          2021-12-31             280.           21093.
#  4 South Dakota          2021-12-31             278.           20014.
#  5 Wyoming               2021-12-31             264.           19979.
#  6 Tennessee             2021-12-31             296.           19783.
#  7 Kentucky              2021-12-31             269.           19173.
#  8 Florida               2021-12-31             287.           19128.
#  9 Utah                  2021-12-31             113.           19088.
# 10 Wisconsin             2021-12-31             190.           19008.
# ... with 41 more rows
```

-- Communicate your methodology, results, and interpretation here -- 

It looks like the list from Question 1 is very different to the above list when calculated per 100,000 people. States such as California and Texas which were at the top of the list in Q1 is not even in the top 10 when calculating per 100,000. This goes to show how population plays a role in determining how affected a place is. 
The list in Q2 shows how states have been affected relatively when their respective population is considered. States such as North Dakota and Alaska may not have had high occurrences of deaths and cases overall, but considering its population, it starts to rank high. The second method seems to be a better indicator of how responsive the state has been to in its efforts to curb the spread of the disease. 

##### Question 3

Now, select a state and calculate the seven-day averages for new cases and deaths per 100,000 people. Once you have calculated the averages, create a visualization using ggplot2 to represent the data. 

```{r p2q3-response}

# Select a state and then filter by state and date range your data from Question 1. Calculate the seven-day average following the same procedure as Part 1. 

# selecting the state of colorado and obtaining sum of total deaths and cases
colorado_df <- combined_df %>% 
  filter(state == 'Colorado', date >= "2020-03-15", date <= "2021-12-31") %>% 
  group_by(state, date) %>% 
  summarise(total_deaths = sum(deaths), 
            total_cases = sum(cases)) %>% 
  mutate(year = year(date))

# using tha above information and after performing a join with its population data - calculating the rolling averages
rolling_colorado_df <- colorado_df %>% 
    left_join(state_pop_df, by = c("state" = "STNAME", "year" = "Year")) %>% 
    mutate(deaths_per_100k = (total_deaths/total_pop)*100000, 
           cases_per_100k = (total_cases/total_pop)*100000,
           deaths_7_day = (rollapply(c(NA, diff(total_deaths)), width = 7, FUN = mean, align = "right", fill=NA)/total_pop)*100000, 
           cases_7_day = (rollapply(c(NA, diff(total_cases)), width = 7, FUN = mean, align = "right", fill = NA)/total_pop)*100000)

# viewing the data
rolling_colorado_df

# setting a coefficient to visualize the graphs and scales better
coeff <- 10

# plotting a line graph of rolling averages of cases vs deaths 
ggplot(rolling_colorado_df, aes(x = date)) +
  geom_line(aes(y = deaths_7_day, color = "Deaths"), size = 1) +  # Raw scores as bar chart
  geom_line(aes(y = cases_7_day/coeff, color = "Cases"), size = 1) +  # Rolling average as line
  labs(x = "Date", color = "Metric") +
  scale_color_manual(values = c("Deaths" = "red", "Cases" = "blue")) +
  theme_bw() + 
  scale_y_continuous(
        name = "Deaths",
        #labels = unit_format(unit = "M", scale = 100e-6), 
        sec.axis = sec_axis(~.*10, name = "Cases")  # Adjust the transformation
    ) +
    ggtitle("Rolling averages of deaths & cases follow a similar trend!")
  
```

```{r p2q3-sample}

# Your transformed data should look similar to the following tibble: 
#
# A tibble: 656 × 9
#     state    date       total_deaths total_cases population deaths_per_100k cases_per_100k deaths_7_day  cases_7_day
#     <chr>    <date>            <dbl>    <dbl>      <dbl>         <dbl>          <dbl>        <dbl>         <dbl>
#  1 Colorado 2020-03-15            2      136      5784308        0.0346          2.35         NA            NA   
#  2 Colorado 2020-03-16            2      161      5784308        0.0346          2.78         NA            NA   
#  3 Colorado 2020-03-17            3      183      5784308        0.0519          3.16         NA            NA   
#  4 Colorado 2020-03-18            3      216      5784308        0.0519          3.73         NA            NA   
#  5 Colorado 2020-03-19            5      278      5784308        0.0864          4.81         NA            NA   
#  6 Colorado 2020-03-20            5      364      5784308        0.0864          6.29         NA            NA   
#  7 Colorado 2020-03-21            6      475      5784308        0.104           8.21         NA            NA   
#  8 Colorado 2020-03-22            7      591      5784308        0.121           10.2       0.0123         1.12
#  9 Colorado 2020-03-23           10      721      5784308        0.173           12.5       0.0198         1.38
# 10 Colorado 2020-03-24           11      912      5784308        0.190           15.8       0.0198         1.80
# … with 646 more rows
```

-- Communicate your methodology, results, and interpretation here -- 

After calculating the rolling averages for deaths and cases for the state of Colorado, the same as been plotted onto a line graph. They seem to follow a similar pattern as expected - when cases go up, deaths also go up. Colorado seems to not have been affected as badly as the other states in the list as per Q2. Relatively, based on it's population, it was not in the top 10. 

##### Question 4

Using the same state, identify the top 5 counties in terms of deaths and cases per 100,000 people. 

```{r p2q4-response}

# Using the same state as Question 2, filter your state and date range from the combined data set from Part 1 and summarize cases and deaths. Produce two lists arranged by deaths and cases. When transforming the data, be sure to include the "fips" column as you will need this to complete Question 5. 

# getting the total deaths and cases arranged by deaths in descending order
county_d_df <- combined_df %>% 
  filter(state == "Colorado", date >= "2020-03-15", date <= "2021-12-20") %>% 
  group_by(county, date) %>% 
  summarise(total_deaths = sum(deaths), 
            total_cases = sum(cases)) %>% 
  arrange(desc(total_deaths), desc(total_cases)) %>% 
  distinct(county, .keep_all = TRUE) %>% 
  mutate(year = year(date))

# using another df to get the fips column
county_fips_df <- combined_df %>% 
  filter(state == "Colorado", date >= "2020-03-15", date <= "2021-12-20") %>% 
  select(county, fips) %>% 
  distinct(county, .keep_all = TRUE)

# joining the above two tables to get all the required columns arranged by deaths 
county_d_df <- county_d_df %>% 
  left_join(county_fips_df, by = "county") %>% 
  select(county, date, fips, everything())

# getting the total deaths and cases arranged by cases in descending order
county_c_df <- combined_df %>% 
  filter(state == "Colorado", date >= "2020-03-15", date <= "2021-12-20") %>% 
  group_by(county, date) %>% 
  summarise(total_deaths = sum(deaths), 
            total_cases = sum(cases)) %>% 
  arrange(desc(total_cases), desc(total_deaths)) %>% 
  distinct(county, .keep_all = TRUE) %>% 
  mutate(year = year(date))

# joining the above two tables to get all the required columns arranged by cases 
county_c_df <- county_c_df %>% 
  left_join(county_fips_df, by = "county") %>% 
  select(county, date, fips, everything())

# getting the population details for each of the counties of Colorado
county_pop_df <- us_population_estimates %>% 
  filter(STNAME == "Colorado") %>% 
  separate(CTYNAME, into = c("county"), sep = " County")

# calculating the deaths and cases per 100k for the df that is arranged by deaths 
county_d_df <- county_d_df %>% 
  left_join(county_pop_df, by = c("county", "year" = "Year")) %>% 
  select(county, date, fips.x, total_deaths, total_cases, year, Estimate) %>% 
  rename("fips" = "fips.x") %>% 
  mutate(deaths_per_100k = (total_deaths/Estimate)*100000, 
         cases_per_100k = (total_cases/Estimate)*100000)

# calculating the deaths and cases per 100k for the df that is arranged by cases 
county_c_df <- county_c_df %>% 
  left_join(county_pop_df, by = c("county", "year" = "Year")) %>% 
  select(county, date, fips.x, total_deaths, total_cases, year, Estimate) %>% 
  rename("fips" = "fips.x") %>% 
  mutate(deaths_per_100k = (total_deaths/Estimate)*100000, 
         cases_per_100k = (total_cases/Estimate)*100000)

# viewing the data 
county_d_df
county_c_df
```

```{r p2q4-sample}

# Your transformed data should be similar to the following tibbles: 
#
# Arranged by deaths: 
# A tibble: 64 × 4
#     county       date       fips    total_deaths   total_cases
#     <chr>       <date>      <chr>       <dbl>         <dbl>
#  1 El Paso    2021-12-20    08041       1355         119772
#  2 Denver     2021-12-20    08031       1065         106747
#  3 Jefferson  2021-12-20    08059       1061         76732
#  4 Adams      2021-12-20    08001       1057         90476
#  5 Arapahoe   2021-12-20    08005       1046         95769
#  6 Pueblo     2021-12-20    08101        643         30739
#  7 Weld       2021-12-20    08123        569         55599
#  8 Mesa       2021-12-20    08077        445         29542
#  9 Larimer    2021-12-20    08069        393         47444
# 10 Douglas    2021-12-20    08035        361         48740
# … with 54 more rows
#
#
# Arranged by cases: 
# A tibble: 64 × 4
#     county       date       fips   total_deaths   total_cases
#     <chr>       <date>      <chr>     <dbl>         <dbl>
#  1 El Paso    2021-12-20    08041     1355         119772
#  2 Denver     2021-12-20    08031     1065         106747
#  3 Arapahoe   2021-12-20    08005     1046          95769
#  4 Adams      2021-12-20    08001     1057          90476
#  5 Jefferson  2021-12-20    08059     1061          76732
#  6 Weld       2021-12-20    08123      569          55599
#  7 Douglas    2021-12-20    08035      361          48740
#  8 Larimer    2021-12-20    08069      393          47444
#  9 Boulder    2021-12-20    08013      323          36754
# 10 Pueblo     2021-12-20    08101      643          30739
# … with 54 more rows
```

-- Communicate your methodology, results, and interpretation here -- 

To get the required tables county wise, I have grouped the combined_df using county and date after adding the necessary filters. For some reason, I could not get the distinct values of county after arranging them when grouping by fips as well, hence I have extracted the fips in a separate df and performed a left join to add the fips column in the end result as well. Then I have preprocessed the us population dataset for the counties of Colorado and used the population to calculate the deaths and cases per 100,000 people. 

##### Question 5

Modify the code below for the map projection to plot county-level deaths and cases per 100,000 people for your state. 

```{r p2q5-response}

# plotting the deaths_per_100k for each county of Colorado using maps
plot_usmap(regions = "county", include="CO", data = county_d_df, values = "deaths_per_100k", color = "blue") +
  scale_fill_continuous(low = "white", high = "red", name = "Deaths per 100,000") +
  ggtitle("Deaths per 100,000 in Colorado Counties")

# plotting the cases_per_100k for each county of Colorado using maps
plot_usmap(regions = "county", include="CO", data = county_c_df, values = "cases_per_100k", color = "blue") +
  scale_fill_continuous(low = "white", high = "blue", name = "Cases per 100,000") +
  ggtitle("Cases per 100,000 in Colorado Counties")

```

```{r p2q5-sample}
# Copy and modify the code below for your state.
#
# plot_usmap arguments: 
#   regions: can be one of ("states", "state", "counties", "county"). The default is "states"
#   include: The regions to include in the resulting map. If regions is "states"/"state", the value can be either a state name, abbreviation or FIPS code. For counties, the FIPS must be provided as there can be multiple counties with the same name. 
#   data: values to plot on the map
#   values: the name of the column that contains the values to be associated with a given region. 
#   color: the map outline color.
#
# Reference the plot_usmap documentation for further information using ?plot_usmap

#plot_usmap(regions = "county", include="CO", data = colorado_county, values = "total_deaths", color = "blue") +
#  scale_fill_continuous(low = "white", high = "blue", name = "Deaths per 100,000")

```

-- Communicate your methodology, results, and interpretation here -- 

It looks like the County Oteri, though its low population, was severely affected both by terms of the total cases and deaths that occurred here. Other counties such as Crowley and Bent were also highly affected given their low population when taking into consideration of the metric per 100,000 people. 

##### Question 6

Finally, select three other states and calculate the seven-day averages for new deaths and cases per 100,000 people for between March 15, 2020, and December 31, 2021. 

```{r p2q6-response}

# selecting the state of California and obtaining sum of total deaths and cases
california_df <- combined_df %>% 
  filter(state == 'California', date >= "2020-03-15", date <= "2021-12-31") %>% 
  group_by(state, date) %>% 
  summarise(total_deaths = sum(deaths), 
            total_cases = sum(cases)) %>% 
  mutate(year = year(date))

# using tha above information and after performing a join with its population data - calculating the rolling averages
rolling_california_df <- california_df %>% 
    left_join(state_pop_df, by = c("state" = "STNAME", "year" = "Year")) %>% 
    mutate(deaths_per_100k = (total_deaths/total_pop)*100000, 
           cases_per_100k = (total_cases/total_pop)*100000,
           deaths_7_day = (rollapply(c(NA, diff(total_deaths)), width = 7, FUN = mean, align = "right", fill=NA)/total_pop)*100000, 
           cases_7_day = (rollapply(c(NA, diff(total_cases)), width = 7, FUN = mean, align = "right", fill = NA)/total_pop)*100000)

# selecting the state of New York and obtaining sum of total deaths and cases
newyork_df <- combined_df %>% 
  filter(state == 'New York', date >= "2020-03-15", date <= "2021-12-31") %>% 
  group_by(state, date) %>% 
  summarise(total_deaths = sum(deaths), 
            total_cases = sum(cases)) %>% 
  mutate(year = year(date))

# using tha above information and after performing a join with its population data - calculating the rolling averages
rolling_newyork_df <- newyork_df %>% 
    left_join(state_pop_df, by = c("state" = "STNAME", "year" = "Year")) %>% 
    mutate(deaths_per_100k = (total_deaths/total_pop)*100000, 
           cases_per_100k = (total_cases/total_pop)*100000,
           deaths_7_day = (rollapply(c(NA, diff(total_deaths)), width = 7, FUN = mean, align = "right", fill=NA)/total_pop)*100000, 
           cases_7_day = (rollapply(c(NA, diff(total_cases)), width = 7, FUN = mean, align = "right", fill = NA)/total_pop)*100000)

# selecting the state of Minnesota and obtaining sum of total deaths and cases
minnesota_df <- combined_df %>% 
  filter(state == 'Minnesota', date >= "2020-03-15", date <= "2021-12-31") %>% 
  group_by(state, date) %>% 
  summarise(total_deaths = sum(deaths), 
            total_cases = sum(cases)) %>% 
  mutate(year = year(date))

# using tha above information and after performing a join with its population data - calculating the rolling averages
rolling_minnesota_df <- minnesota_df %>% 
    left_join(state_pop_df, by = c("state" = "STNAME", "year" = "Year")) %>% 
    mutate(deaths_per_100k = (total_deaths/total_pop)*100000, 
           cases_per_100k = (total_cases/total_pop)*100000,
           deaths_7_day = (rollapply(c(NA, diff(total_deaths)), width = 7, FUN = mean, align = "right", fill=NA)/total_pop)*100000, 
           cases_7_day = (rollapply(c(NA, diff(total_cases)), width = 7, FUN = mean, align = "right", fill = NA)/total_pop)*100000)

# viewing the data
rolling_colorado_df
rolling_california_df
rolling_newyork_df
rolling_minnesota_df
```

-- Communicate your methodology, results, and interpretation here -- 

Following the same method that was followed for the state of Colorado in Question 3. The other three states that are included are California, New York and Minnesota. 
Picked California and New York for their huge population and Colorado and Minnesota for their relatively lesser population. This is done to see how population plays a role in rolling 7 day average metrics for deaths and cases. 

##### Question 7

Create a visualization comparing the seven-day averages for new deaths and cases per 100,000 people for the four states you selected. 

```{r p2q7-response}

# combining the previous dataframes for each state into one df
combined_rolling_state_df <- rbind(rolling_colorado_df, rolling_california_df, rolling_newyork_df, rolling_minnesota_df)

# visualizing the rolling 7 days average for deaths for each state
ggplot(combined_rolling_state_df, aes(x = date, y = deaths_7_day, color = fct_reorder2(state, date, deaths_7_day))) +
  geom_line() +
  labs(x = "Date", y = "Deaths (7-day average)", color = "State") +
  ggtitle("Deaths (7-day average) by State") +
  theme_minimal()

# visualizing the rolling 7 days average for cases for each state
ggplot(combined_rolling_state_df, aes(x = date, y = cases_7_day, color = fct_reorder2(state, date, cases_7_day))) +
  geom_line() +
  labs(x = "Date", y = "Cases (7-day average)", color = "State") +
  ggtitle("Cases (7-day average) by State") +
  theme_minimal()

```

-- Communicate your methodology, results, and interpretation here -- 

It looks like at the beginning, New York had the highest number of deaths per 100,000 people when compared to the other 3 states and Minnesota was leading at the end. 
In terms of cases per 100,000 people, New York was leading at the beginning and at the end. This is mostly due to the high population that is present in this state. 
fct_reorder2 is used so that it becomes easier to visualize which state is at the top and also align the labels accordingly. 

#### Part 3 - Global Comparison

```{r import-csse}
# Import global COVID-19 statistics aggregated by the Center for Systems Science and Engineering (CSSE) at Johns Hopkins University.
# Import global population estimates from the World Bank.

csse_global_deaths <- read_csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv")
csse_global_cases <- read_csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv")
csse_us_deaths <- read_csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_US.csv")
csse_us_cases <- read_csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_US.csv")

globabl_population_estimates <- read_csv("global_population_estimates.csv")
```

##### Question 1

Using the state you selected in Part 2 Question 2 compare the daily number of cases and deaths reported from the CSSE and NY Times. 

```{r p3q1-response}

# To compare your state data between the two data sets, you will first need to tidy the US CSSE death and cases data.
# Hint: Review the documentation for pivot_longer().

# obtaining the cases data from csse data
csse_colorado_c_df <- csse_us_cases %>% 
  pivot_longer(cols = -c(1:11), names_to = "year", values_to = "cases") %>% 
  filter(Province_State == "Colorado") %>% 
  mutate(date = as.Date(year, format = "%m/%d/%y")) %>% 
  select(FIPS, Admin2, Province_State, date, cases) %>% 
  rename("fips" = "FIPS", "county" = "Admin2", "state" = "Province_State") %>% 
  filter(date >= "2020-03-15", date <= "2021-12-31")

# obtaining the deaths data from csse data
csse_colorado_d_df <- csse_us_deaths %>% 
  pivot_longer(cols = -c(1:11), names_to = "year", values_to = "deaths") %>% 
  filter(Province_State == "Colorado") %>% 
  mutate(date = as.Date(year, format = "%m/%d/%y")) %>% 
  select(FIPS, Admin2, Province_State, date, deaths) %>% 
  rename("fips" = "FIPS", "county" = "Admin2", "state" = "Province_State") %>% 
  filter(date >= "2020-03-15", date <= "2021-12-31")

# Once you have tidied your data, join the two CSSE US data sets to include cases and deaths in one table. 

# performing a join to combine the above two datasets
csse_colorado_df <- csse_colorado_c_df %>% 
  left_join(csse_colorado_d_df)

# summing up cases and deaths on a state and date level
csse_colorado_total_df <- csse_colorado_df %>% 
  group_by(state, date) %>% 
  summarise(CSSE_cases = sum(cases),
            CSSE_deaths = sum(deaths))

# renaming columns for convenience 
colorado_df <- colorado_df %>% 
  rename("NY_Times_cases" = "total_cases", "NY_Times_deaths" = "total_deaths")

final_colorado_df <- colorado_df %>% 
  full_join(csse_colorado_total_df)

# Finally, create two visualizations with one plotting the CSSE and NY Times cases and the other plotting the CSEE and NY Times deaths.

# plotting a line graph for comparing the NY Times vs CSSE cases data 
ggplot(final_colorado_df, aes(x = date)) +
  geom_line(aes(y = NY_Times_cases, color = "NY Times Cases"), size = 1) +
  geom_line(aes(y = CSSE_cases, color = "CSSE Cases"), size = 1) +
  labs(x = "Date", y = "Cases Reported", color = "Metric") +
  scale_color_manual(values = c("CSSE Cases" = "orange", "NY Times Cases" = "blue")) +
  theme_bw() +
  ggtitle("The NY Times and CSSE data have reported the same data on no of cases!")

# plotting a line graph for comparing the NY Times vs CSSE deaths data 
ggplot(final_colorado_df, aes(x = date)) +
  geom_line(aes(y = NY_Times_deaths, color = "NY Times Deaths"), size = 1) +
  geom_line(aes(y = CSSE_deaths, color = "CSSE Deaths"), size = 1) +
  labs(x = "Date", y = "Deaths Reported", color = "Metric") +
  scale_color_manual(values = c("CSSE Deaths" = "orange", "NY Times Deaths" = "blue")) +
  theme_bw() +
  ggtitle("CSSE reported slightly more deaths at certain times!")

```

```{r p3q1-sample}

# Your tidied CSSE data for your selected state should look similar to the following tibble:
#
# A tibble: 43,362 × 6
#     fips  county  state        date      cases  deaths
#     <dbl> <chr>   <chr>       <date>     <dbl>  <dbl>
#  1  8001  Adams  Colorado   2020-03-15     6      0
#  2  8001  Adams  Colorado   2020-03-16     8      0
#  3  8001  Adams  Colorado   2020-03-17    10      0
#  4  8001  Adams  Colorado   2020-03-18    10      0
#  5  8001  Adams  Colorado   2020-03-19    10      0
#  6  8001  Adams  Colorado   2020-03-20    12      0
#  7  8001  Adams  Colorado   2020-03-21    14      0
#  8  8001  Adams  Colorado   2020-03-22    18      0
#  9  8001  Adams  Colorado   2020-03-23    25      0
# 10  8001  Adams  Colorado   2020-03-24    27      0
# … with 43,352 more rows

```

-- Communicate your methodology, results, and interpretation here -- 

When comparing the NY Times and CSSE data, it looks like both reported more or less the same numbers wrt the number of cases during this duration. But when it comes to the deaths, CSSE has reported slightly more than NY Times for the Colorado State, especially during the beginning, end of 2020 and towards the end of 2021. These were also the moments when the number of deaths and cases were increasing according to the data reported by both of them. During such times, it can be expected that the data given by both may not be similar due to some underestimation and overestimation that can easily occur. 

##### Question 2 

Now that you have verified the data reported from the CSSE and NY Times are similar, combine the global and US CSSE data sets and identify the top 10 countries in terms of deaths and cases per 100,000 people between March 15, 2020, and December 31, 2021.

```{r p3q2-response}

# First, combine and tidy the CSSE death and cases data sets. You may wish to keep the two sets separate.
# Then, tidy the global population estimates. While tidying your data, remember to include columns that you will be able to use when joining the COVID-19 data. 
# You will notice that the population estimates data does not include every country reported in the CSSE data. When calculating statistics per 100,000 people, you will need to filter the CSSE data to only include countries that you have population estimates for. 

# tidying up the global cases dataset
global_cases_df <- csse_global_cases %>% 
  pivot_longer(cols = -c(1:4), names_to = "year", values_to = "cases") %>% 
  mutate(date = as.Date(year, format = "%m/%d/%y")) %>% 
  rename("country" = "Country/Region") %>% 
  filter(date >= "2020-03-15", date <= "2021-12-31") %>% 
  mutate(year = year(date))

# tidying up the global deaths dataset
global_deaths_df <- csse_global_deaths %>% 
  pivot_longer(cols = -c(1:4), names_to = "year", values_to = "deaths") %>% 
  mutate(date = as.Date(year, format = "%m/%d/%y")) %>% 
  rename("country" = "Country/Region") %>% 
  filter(date >= "2020-03-15", date <= "2021-12-31") %>% 
  mutate(year = year(date))

# tidying up the global population dataset
global_pop_df <- globabl_population_estimates %>% 
  pivot_longer(cols = -c(1:4), names_to = "year", values_to = "population") %>% 
  separate(year, into = c("year"), sep = " ") %>% 
  rename("country" = "Country Name") %>% 
  mutate(year = as.numeric(year), 
         population = as.numeric(population))

# performing the joins needed to combine above three datasets and caluclating the cases and deaths per 100,000 people
global_combined_df <- global_cases_df %>% 
  full_join(global_deaths_df) %>% 
  inner_join(global_pop_df) %>% 
  select(country, Lat, Long, year, date, cases, deaths, population) %>% 
  mutate(deaths_per_100k = (deaths/population)*100000, 
         cases_per_100k = (cases/population)*100000)

# viewing the top 10 countries with highest deaths per 100,000 people
top_10_deaths_global_df <- global_combined_df %>% 
  arrange(desc(deaths_per_100k)) %>% 
  distinct(country, .keep_all = TRUE) %>% 
  head(n = 10)

# viewing the top 10 countries with highest cases per 100,000 people
top_10_cases_global_df <- global_combined_df %>% 
  arrange(desc(cases_per_100k)) %>% 
  distinct(country, .keep_all = TRUE) %>% 
  head(n = 10)

top_10_deaths_global_df
top_10_cases_global_df
```

-- Communicate your methodology, results, and interpretation here -- 

When it comes to the highest deaths per 100,000 people, it looks like a lot of European countries suffered the most. A lot of these countries like Bulgaria, Hungary, Moldova, Croatia are on the top 10. 
This is the same case when it comes to cases per 100,000 people too. Countries in the top 10 include United Kingdom, Serbia, San Marino. It is also surprising to see Seychelles - which is a remote island off the coast of East Africa. 

##### Question 3

Construct a visualization plotting the 10 countries in terms of deaths and cases per 100,000 people between March 15, 2020, and December 31, 2021. In designing your visualization keep the number of data you will be plotting in mind. You may wish to create two separate visualizations, one for deaths and another for cases. 

```{r p3q3-response}

# Creating a separate dataset for visualizing cases per 100,000 people
top_10_cases_visual_df <- global_combined_df %>% 
    filter(country %in% top_10_cases_global_df$country) %>% 
    group_by(country, date) %>% 
    summarise(cases_per_100k = sum(cases_per_100k))

# Creating a separate dataset for visualizing deaths per 100,000 people
top_10_deaths_visual_df <- global_combined_df %>% 
    filter(country %in% top_10_deaths_global_df$country) %>% 
    group_by(country, date) %>% 
    summarise(deaths_per_100k = sum(deaths_per_100k))

# visualizing cases for the top 10 countries with highest cases per 100,000 people
ggplot(top_10_cases_visual_df, aes(x = date, y = cases_per_100k, color = fct_reorder2(country, date, cases_per_100k))) +
  geom_line(size = 0.5) +
  labs(x = "Date", y = "Cases per 100,000 people", color = "Country") +
  ggtitle("COVID-19 Cases per 100,000 People") +
  theme_bw()

# visualizing deaths for the top 10 countries with highest deaths per 100,000 people
ggplot(top_10_deaths_visual_df, aes(x = date, y = deaths_per_100k, color = fct_reorder2(country, date, deaths_per_100k))) +
  geom_line(size = 0.5) +
  labs(x = "Date", y = "Deaths per 100,000 people", color = "Country") +
  ggtitle("COVID-19 Deaths per 100,000 People") +
  theme_bw()
```

-- Communicate your methodology, results, and interpretation here -- 

In terms of cases, Andorra leads the list. It ranked consistently at the top throughout the entire duration of 2 years as well. 
But when it comes to death per 100,000 people, Peru is ranked not just at the top but has a significant difference when compared to the other countries. This could have been due to several factors such as non-speedy distribution of vaccines and overcrowding of certain areas with a lot of homes. 

##### Question 4

Finally, select four countries from one continent and create visualizations for the daily number of confirmed cases per 100,000 and the daily number of deaths per 100,000 people between March 15, 2020, and December 31, 2021. 

```{r p3q4-response}

asia_df <- global_combined_df %>% 
  mutate(daily_cases = c(NA, diff(cases)), 
         daily_deaths = c(NA, diff(deaths)), 
         daily_cases_per_100k = (daily_cases/population)*100000, 
         daily_deaths_per_100k = (daily_deaths/population)*100000) %>% 
  filter(country %in% c("India", "Singapore", "Japan", "Malaysia"), daily_cases_per_100k >= 0, daily_deaths_per_100k >= 0)

ggplot(asia_df, aes(x = date, y = daily_cases_per_100k, fill = fct_reorder2(country, date, daily_cases_per_100k))) +
  geom_area() +
  labs(x = "Date", y = "Daily Cases per 100,000", fill = "Country") +
  ggtitle("Daily Cases per 100,000 People") +
  theme_minimal()

ggplot(asia_df, aes(x = date, y = daily_deaths_per_100k, fill = fct_reorder2(country, date, daily_deaths_per_100k))) +
  geom_area() +
  labs(x = "Date", y = "Daily Deaths per 100,000", fill = "Country") +
  ggtitle("Daily Deaths per 100,000 People") +
  theme_minimal()
```

-- Communicate your methodology, results, and interpretation here -- 

I've used an area chart to indicate which country has the highest daily deaths and cases per 100,000 people. Countries from Asia - India, Malaysia, Singapore and Japan are considered. When it comes to daily cases, it looks like Malaysia is at the top throughout mostly. Especially towards the very end. This could be the case that this country has a low population and a high case rate for its population. 
Even for deaths, Malaysia is at the top with India having some significant numbers closer to Aug of 2021. 